Do return series have power-law tails with the same index?

### 2017-04-17 papers/FX/draft.tex

\documentclass[11pt,a4]{amsart}
%\usepackage{amsmath}
%\usepackage{amsthm}
%\usepackage{enumerate}
\usepackage[bookmarks=true]{hyperref}
%% \usepackage{bookmark}
\usepackage{graphicx}
%\usepackage{multirow}
%\usepackage{adjustbox}
%\usepackage{wrapfig}
%\usepackage{subcaption}
\usepackage{color}
\usepackage[dvipsnames]{xcolor}

@@

\subsection{Hill estimates of lower-tail scale parameters}\label{sec:HillScaleEstimates}
We assume the pure Pareto model \eqref{eq:3} with scale parameter $K>0$.
Hill \cite{hill1975simple} proposed  the maximum-likelihood estimator of $K$
derived from the joint \ds\ of $k$ upper order statistics in the sample; cf.
Embrechts et al. \cite{embrechts:klueppelberg:mikosch:1997}, p.~334.
It is given by
\beao
\hat K_k = \left({k \over n}\right)^{1/\hat\alpha} X_{(n-k)}\,.
\eeao
Using the asymptotic normality property of upper order statistics
(cf. de Haan and Ferreira~\cite{haan:ferreira:2006}, theorem 2.2.1), one can show
\begin{eqnarray*}
  \sqrt k (\hat K_k - K) &\overset{d}{\to}& N\left(
    0, \left( {K \over \alpha}  \right)^2
  \right) \\
  \sqrt k (\hat K_k^\alpha - K^\alpha) &\overset{d}{\to}& N\left(
    0, \left( {K^{2\alpha}}  \right)^2
  \right)
\end{eqnarray*}
where the tail index $\alpha$ is regarded as known. From the above
asymptotic normality property, confidence bands of the estimate of $K$
and $K^\alpha$ can be constructed. 
Estimates of $K$ of the ``Energy'', ``Consumer Staples'' and
``Information Technology'' sectors of the S\&P 500 index are computed
using this method. The results are shown in
Figure~\ref{fig:sectors_parameters}.
\begin{figure}[htb!]
  \centering
  \begin{minipage}{0.33\linewidth}
    \includegraphics[width=\textwidth]
    {Energy_K.pdf}
  \end{minipage}\hfill
  \begin{minipage}{0.33\linewidth}
    \includegraphics[width=\textwidth]
    {Consumer_Staples_K.pdf}
  \end{minipage}\hfill
  \begin{minipage}{0.33\linewidth}
    \includegraphics[width=\textwidth]
    {Information_Technology_K.pdf}
  \end{minipage}
  \begin{minipage}{0.33\linewidth}
    \includegraphics[width=\textwidth]
    {Energy_scale.pdf}
  \end{minipage}\hfill
  \begin{minipage}{0.33\linewidth}
    \includegraphics[width=\textwidth]
    {Consumer_Staples_scale.pdf}
  \end{minipage}\hfill
  \begin{minipage}{0.33\linewidth}
    \includegraphics[width=\textwidth]
    {Information_Technology_scale.pdf}
  \end{minipage}
  \caption{\small Estimates of $K$ and $K^\alpha$ of stocks in sectors
    of the S\&P 500 index. The points are the estimated values; the bars are
    the $2\sigma$ confidence intervals. The confidence bands of the
    corresponding Hill estimates $\hat \alpha$ of these sectors are
    shown in figure~\ref{fig:1}. Note $\log_{10}(K^\alpha)$ is shown
    instead of $K^\alpha$.
  }
  \label{fig:sectors_parameters}
\end{figure}
One can see that $K$ generally takes a rather small value. For the
more volatile sectors of ``Energy'' and ``Information Technology'',
the average value of $K$ is around 0.01, while for the more stable
sector of ``Consumer Staples'', the average value is around 0.005. Due
to the smallness of $K$, mild variations of $\alpha$ would lead to
huge variations of $K^\alpha$, as shown in the 2nd column of
figure~\ref{fig:sectors_parameters}.

@@

argue this can be the result of slow convergence of the distribution
of the test statistic to its asymptotic distribution. We elaborate on
this in the section~\ref{sec:Hoga}.

Nevertheless, one should to bear in mind that the weak convergence
\eqref{eq:x3} is valid only on condition that both the $X_i$ and $Y_i$
series are independent of each other, which is generally untrue for
two return series in the same market.

%% By the design of Hoga's test, the null hypothesis can be
%% rejected due to changed scale parameters instead of changed tail
%% indices. Thus it is useful to apply the Hill scale estimator to the
%% same stock return series.

@@

To the best of our knowledge, there do not exist statistical tests for comparing the 
tail indices or scale parameters of two or more dependent \ts\ in  a direct way. {\red construct a test for this}
We borrow a test from a recent paper by Hoga \cite{hoga:2016}. 
This test has been developed for a different kind of problem. Given a strictly stationary \ts\ $X_1,\ldots,X_n$
with a marginal \ds\ $F$ with right power-law tail, the goal is to test whether there is a structural break 
of the {\em extreme quantiles} $F^{-1}(1-p)$ for values $p$ very close to zero. If the tail index
{\em or} the scale parameter in a \ds\ of type \eqref{eq:3} change inside a sample, then it is likely
that the extreme quantiles change as well. We will test for a change of tail index or scale parameter 
in this indirect way.

@@

\subsection{An utility based argument for equal tail indices}\label{sec:3}
In this section we give an argument based on economic theory which  favors equality of  tail indices for return series.
We follow an approach by Routledge and Zin \cite{routledge2010generalized} who introduced the notion of
Generalized Disappointment Aversion (GDA). We consider the risky payoff $C$ of an investor and assume that
it has a continuous \ds\ on $(0,\infty)$ with \df\ $F_C$. Let $u$ be a utility \fct\ assumed to be increasing and concave on $(0,\infty)$.
Following Routledge and Zin \cite{routledge2010generalized},
the utility of an agent with GDA preferences is given by
\beao% \label{eq:xxie0}
  \wt u&=& \E [u(C)] - b\, \int_{0}^{\delta v}
  \big[ u(\delta \,v) - u(x) \big] F_C(dx)\nonumber\\&=& \E u(C) - b \E\big[\big(u(\delta \,v) - u(C)\big)\1( C < \delta \,v)\big]\,,
\eeao
where $\delta $ and $v$ are positive constants, and $b\ge 0$. {\red
  meaning of these constants?}
{\blue 
 Here $v$ can be thought of as the certainty payoff equivalent to the
 risky payoff $C$; $\delta$ tunes the threshold of {\em
   disappointment} in proportion to $v$; $b$ determines the extra
 weight given to the expected return of $C$ when $C$ is below the
 disappoinment threshold $\delta v$.
 If $b=0$, preferences are the classical expected utility. If
 $\delta=1$ and $b>0$ preferences follow Gul's \cite{gul:1991} 
disappointment aversion which were generalized by Routledge and Zin
\cite{routledge2010generalized}.
}
An agent guided by the utility function $u$ will seek to maximize the
\fct al $\wt u$.
Routledge and Zin assumed a power-law utility function 
\beam\label{eq:u}
u(x)=\frac{1}{1 - \gamma}\,x^{1 - \gamma}\,,\quad \gamma>0\,. 
\eeam
In practice it is more often to observe $\gamma > 1$ {\blue Let's
consult Casper for references.} Therefore we adopt a more convenient
notation in this case
\beao
u(x)=-\frac{1}{\xi}\,x^{-\xi}\,,\quad \xi>0\,. 
\eeao

@@

\subsection{Pareto-distributed returns}\label{sec:pareto_tail}
We consider the case when $X$ has a two-sided Pareto \ds\ given by
\beam\label{eq:pareto}
  F_X(x) = \left\{
  \begin{array}{ll}
    p \left(
    {K \over K - x}
    \right)^\alpha & x \leq 0 \,,\\
    1 - (1 - p) \left(
    {K' \over K' + x}
    \right)^\beta & x > 0\,,
  \end{array}
  \right.
\eeam
where 
$\alpha, \beta > 1$, $0 < p < 1$ and $K,K' > 0$. Then we have
\beam\label{eq:xxie1.0}
 \wt u(F_X,\phi)\nonumber
  &=&
  \alpha\, K^\alpha\,  p\,
  \int_{-\infty}^0
  u\big( (1 - \phi) \ex^r + \phi \ex^x \big)\,
  {1 + b \over (K - x)^{\alpha + 1}} \,dx\\
    &&+
  \beta \,(K')^\beta\, (1 - p)
  \int_{0}^\infty
  u\big( (1 - \phi) \ex^r + \phi\, \ex^x \big)\,
  {1 + b \1(x < q) \over (K' + x)^{\beta + 1}} \,dx 
  - b \,u(\delta v) \,F_X(q) \,.\label{eq:pp}
\eeam
    {\red Is this worth a theorem?}
    {\blue No, let's change it to a remark.}
\begin{remark}\label{thrm:I}
 Assume the two-sided Pareto model \eqref{eq:pp} and that the utility
 \fct\ $u$ is increasing and differentiable. Then $\frac{\pd \wt
   u_{\rm max}}{\pd \alpha} > 0$ and $\frac{\pd \wt u_{\rm max}}{\pd
   K} < 0$.
\end{remark}
The proof is given in Appendix~\ref{sec:thrmI_proof}. 
\par
We illustrate the results of remark~\ref{thrm:I} in Figure~\ref{fig:preference_pareto} for the power-utility
\fct\ \eqref{eq:u}.
For this reason, we replace $\phi$ in the first integral in \eqref{eq:pp} by the optimal $\hat\phi$. The resulting
quantity is denoted by $I(\alpha, K)$.
\par
We conclude that
$\wt u_{\rm max}$ increases with $\alpha$ and decreases with $K$. Therefore there is a curve of
equal preference on the $(\alpha, K)$-plane. Moving along this
curve in the direction of increasing $\alpha$, one expects the values
of $K$ to increase too, i.e., the estimated values of $\alpha$ and $K$
should appear positively dependent. This is indeed the case for some
real return data, e.g. the ``Energy'', ``Consumer Staples'' and
``Information Technology'' sectors of the S\&P 500 index; see 
Figure \ref{fig:sectors_parameters}.
\par
The value of $\hat\phi$ can be calculated
by numerical integration and optimization with respect to $\phi$ for given values of $K, K', \alpha, \beta$. 
This is shown in 
Figure \ref{fig:phi_hat_pareto} for the power utility \fct\ \eqref{eq:u}.
\begin{figure}
  \begin{minipage}{0.5\linewidth}
    \includegraphics[width=\textwidth]{phi_hat_pareto5e-1.pdf}    
  \end{minipage}\hfill
  \begin{minipage}{0.5\linewidth}
    \includegraphics[width=\textwidth]{phi_hat_pareto4.pdf}
  \end{minipage}
  \caption{The graphs show $\hat\phi$, the optimal equity allocation
    as a \fct\ of $\alpha$ and $K$ in the two-sided Pareto model
    \eqref{eq:pareto} with $K=K'$, $\alpha=\beta$. We choose
    the utility $u$ from \eqref{eq:u}
    function $u(x) = -x^{-\xi}/\xi$ for $\xi = 1/2$ (left) and $\xi = 4$
    (right). {\rm change the $u(C)$ to $u(x)$ everywhere also in other
      figures, and use the same structure of the formulas.}
  }
  \label{fig:phi_hat_pareto}
\end{figure}

\begin{minipage}{0.5\linewidth}
  \includegraphics[width=\textwidth]{power_utilities.png}
\end{minipage}\hfill
\begin{minipage}{0.42\textwidth}{\red change formulas in graph and $C$ to $x$.}
  In this graph we show two power-utility functions $u$. We see that $u(x)$ grows slower
  and saturates later for small $\xi$ than for large $\xi$. The latter case
  represents an agent who is more tolerant about low
  consumption and seeks wealth more aggressively. In other
  words, he is less risk-averse than one with a larger $\xi$.
  Such an agent will therefore invest more heavily in equity, as shown in Figure~\ref{fig:phi_hat_pareto}.
\end{minipage}

@@

\subsection{Student's t-distributed equity return}
It is a common practice to use Student's t-distribution to model the
stationary distribution of equity returns. So it is of interest to
find out what implications this distribution has when it is combined
with the PDA preference. Formally we assume
\[
f(x; \nu) = c(\nu) \left(
  1 + {x^2 \over \nu}
\right)^{-(\nu + 1)/2}
\]
where $\nu > 1$ and
\[
c(\nu) = {
  \Gamma({\nu + 1 \over 2})
  \over
  \Gamma(\nu/2) \sqrt{\nu \pi}
}
\]
We can write $\mathcal U(F, \phi)$ as
\begin{eqnarray}
  && \mathcal U(F, \phi) \nonumber \\
  &=&
  \int_{0}^{\infty}
  \left[
    u(C(x, \phi)) + (1 + b)u(C(-x, \phi))
  \right]
  f(x, \nu) dx \nonumber \\
  &&
  + b \int_{0}^{q}
  u(C(x, \phi))
  f(x, \nu) dx - b u(C(q, \phi)) F(q, \nu)
  \label{eq:t1}
\end{eqnarray}
where $C(\cdot)$ is defined in \eqref{eq:xxie1}.

@@

First notice that when $b = 0$, $\mathcal U(F, \phi)$ reduces to
$\E u(C(X, \phi))$. Only the first integral of \eqref{eq:t1} remains.
Its integrand becomes $\ln[C(x, \phi) C(-x, \phi)]$.
Now observe $\ln(C(x, \phi)C(-x, \phi))$ is an increasing function of
$x$:
\begin{eqnarray*}
  {\pd C(x, \phi) C(-x, \phi) \over \pd x}
  &=&
  \phi e^r (1 - \phi) (e^x - e^{-x}) > 0
\end{eqnarray*}
Thus when $b = 0$, by 1st case of remark \ref{remark:I} and lemma
\ref{lemma:II},
${\pd \mathcal U(F, \phi)) \over \pd \nu} < 0$ for all $\phi$. Hence all
investors guided by this utility function would be seeking the
smallest $\nu$, i.e. the heaviest tail in the market. Here note that
Student's t is a symmetric distribution; a heavier tail implies not
only greater down-side risk but also greater up-side potential.

In fact, even when $b > 0$, the same monotonicity property holds. This
is revealed by numerically maximizing $\mathcal U(F, \phi)$ with
respect to $\phi$ for a sequence of selected values of $\nu$. The
results are shown in figure~\ref{fig:phi_hat_U}.
\begin{figure}[htb!]
  % \begin{subfigure}[b]{0.5\linewidth}
  %   \includegraphics[width=\textwidth]{phi_hat_b_t_log.pdf}
  %   \caption{$\hat\phi(\nu)$ of log-utility function}
  %   \label{fig:phi_hat_t_log}
  % \end{subfigure}
  % \begin{subfigure}[b]{0.5\linewidth}
  %   \includegraphics[width=\textwidth]{U_b_t_log.pdf}
  %   \caption{$G(\nu)$ of log-utility function}
  %   \label{fig:U_b_t_log}
  % \end{subfigure}
  % \begin{subfigure}[b]{0.5\linewidth}
  %   \includegraphics[width=\textwidth]{phi_hat_b_t_power.pdf}
  %   \caption{$\hat\phi(\nu)$ of power-utility function}
  %   \label{fig:phi_hat_t_power}
  % \end{subfigure}
  % \begin{subfigure}[b]{0.5\linewidth}
  %   \includegraphics[width=\textwidth]{U_b_t_power.pdf}
  %   \caption{$G(\nu)$ of power-utility function}
  %   \label{fig:U_b_t_power}
  % \end{subfigure}
  \begin{minipage}{0.5\linewidth}
    \includegraphics[width=\textwidth]{phi_hat_b_t_log.pdf}
  \end{minipage}\hfill
  \begin{minipage}{0.5\linewidth}
    \includegraphics[width=\textwidth]{U_b_t_log.pdf}
  \end{minipage}
  \begin{minipage}{0.5\linewidth}
    \includegraphics[width=\textwidth]{phi_hat_b_t_power.pdf}
  \end{minipage}\hfill
  \begin{minipage}{0.5\linewidth}
    \includegraphics[width=\textwidth]{U_b_t_power.pdf}
  \end{minipage}
  \caption{$\hat\phi$ and $G(\nu)$
    \label{fig:phi_hat_U}
  }
\end{figure}

%% However, this is not necessarily the case when $b > 0$. For example,
%% consider the case when $q = 0$ and $\hat\phi = 1$. This may be
%% viewed an approximation to the situation where the interest rate $r$ is
%% vanishingly low while the investor is more conservative than
%% greedy. In this case we have
%% \begin{eqnarray*}
%%   \mathcal U(F, \phi) = \int_{0}^{\infty} e^{-b x} f(x, \nu) dx
%% \end{eqnarray*}
%% Thus by the 1st case of lemma \ref{lemma:I} and lemma \ref{lemma:II},
%% $\opd{\nu}\mathcal U(F, \phi) > 0$. In this extreme case, the investor
%% would rather seek the largest $\nu$ or in other words, the lightest
%% tail in the market.

\subsubsection{power-utility function}
In this section we consider the case when the utility function
$u(\cdot)$ takes the form
\[
u(C) = -{1 \over \xi} C^{-\xi}
\]
One can see from figure \ref{fig:phi_hat_U_power} that both $\hat
\phi$ and $G(\nu)$ are monotonically decreasing with $\nu$ as in the
case of a log-utility function.
\begin{figure}[htb!]
  \begin{minipage}{0.5\linewidth}
    \includegraphics[width=\textwidth]{phi_hat_b_t_power.pdf}
  \end{minipage}\hfill
  \begin{minipage}{0.5\linewidth}
    \includegraphics[width=\textwidth]{U_b_t_power.pdf}
  \end{minipage}
  \caption{$\hat\phi$ and $G(\nu)$ when the utility
    function is $u(C) = -{2 \over \sqrt{C}}$
  }
  \label{fig:phi_hat_U_power}
\end{figure}

>>>>

\documentclass[11pt,a4]{amsart}
%\usepackage{amsmath}
%\usepackage{amsthm}
%\usepackage{enumerate}
% \usepackage[bookmarks=true]{hyperref}
%% \usepackage{bookmark}
\usepackage{graphicx}
%\usepackage{multirow}
%\usepackage{adjustbox}
%\usepackage{wrapfig}
%\usepackage{subcaption}
\usepackage{color}
\usepackage[dvipsnames]{xcolor}

@@

\subsection{Hill estimates of lower-tail scale parameters}\label{sec:HillScaleEstimates}
We assume the pure Pareto model \eqref{eq:3} with scale parameter $K>0$.
Hill \cite{hill1975simple} proposed  the maximum-likelihood estimator of $K$
derived from the joint \ds\ of $k$ upper order statistics in the sample; cf.
Embrechts et al. \cite{embrechts:klueppelberg:mikosch:1997}, p.~334.
It is given by
\beao
\hat K_k = \left({k \over n}\right)^{1/\hat\alpha} X_{(n-k)}\,.
\eeao
{\blue
Using the asymptotic normality property of upper order statistics
(cf. de Haan and Ferreira~\cite{haan:ferreira:2006}, theorem 2.2.1), one can show
\begin{eqnarray*}
  \sqrt k (\hat K_k - K) &\overset{d}{\to}& N\left(
    0, \left( {K \over \alpha}  \right)^2
  \right) \\
  \sqrt k (\hat K_k^\alpha - K^\alpha) &\overset{d}{\to}& N\left(
    0, \left( {K^{2\alpha}}  \right)^2
  \right)
\end{eqnarray*}
where the tail index $\alpha$ is regarded as known. From the above
asymptotic normality property, confidence bands of the estimate of $K$
and $K^\alpha$ can be constructed. 
Estimates of $K$ of the ``Energy'', ``Consumer Staples'' and
``Information Technology'' sectors of the S\&P 500 index are computed
using this method. The results are shown in
Figure~\ref{fig:sectors_parameters}.
\begin{figure}[htb!]
  \centering
  \begin{minipage}{0.33\linewidth}
    \includegraphics[width=\textwidth]
    {Energy_K.pdf}
  \end{minipage}\hfill
  \begin{minipage}{0.33\linewidth}
    \includegraphics[width=\textwidth]
    {Consumer_Staples_K.pdf}
  \end{minipage}\hfill
  \begin{minipage}{0.33\linewidth}
    \includegraphics[width=\textwidth]
    {Information_Technology_K.pdf}
  \end{minipage}
  \begin{minipage}{0.33\linewidth}
    \includegraphics[width=\textwidth]
    {Energy_scale.pdf}
  \end{minipage}\hfill
  \begin{minipage}{0.33\linewidth}
    \includegraphics[width=\textwidth]
    {Consumer_Staples_scale.pdf}
  \end{minipage}\hfill
  \begin{minipage}{0.33\linewidth}
    \includegraphics[width=\textwidth]
    {Information_Technology_scale.pdf}
  \end{minipage}
  \caption{\small Estimates of $K$ and $K^\alpha$ of stocks in sectors
    of the S\&P 500 index. The points are the estimated values; the bars are
    the $2\sigma$ confidence intervals. The confidence bands of the
    corresponding Hill estimates $\hat \alpha$ of these sectors are
    shown in figure~\ref{fig:1}. Note $\log_{10}(K^\alpha)$ is shown
    instead of $K^\alpha$.
  }
  \label{fig:sectors_parameters}
\end{figure}
One can see that $K$ generally takes a rather small value. For the
more volatile sectors of ``Energy'' and ``Information Technology'',
the average value of $K$ is around 0.01, while for the more stable
sector of ``Consumer Staples'', the average value is around 0.005. Due
to the smallness of $K$, mild variations of $\alpha$ would lead to
huge variations of $K^\alpha$, as shown in the 2nd column of
figure~\ref{fig:sectors_parameters}.

@@

argue this lack of power of the latter can be the result of slow
convergence of the distribution of its test statistic to the
asymptotic distribution. We elaborate on this in section~\ref{sec:Hoga}.

Nevertheless, one should bear in mind that the weak convergence
\eqref{eq:x3} is valid only on condition that both the $X_i$ and $Y_i$
series are independent of each other, which is generally untrue for
two return series in the same market.
}

@@

To the best of our knowledge, there do not exist published statistical
tests for comparing the tail indices or scale parameters of two or
more dependent \ts\ in a direct way. In addition to the test described
in the previous section, we borrow a test from
a recent paper by Hoga \cite{hoga:2016}. This test has been developed
for a different kind of problem. Given a strictly stationary
\ts\ $X_1,\ldots,X_n$ with a marginal \ds\ $F$ with right power-law
tail, the goal is to test whether there is a structural break of the
{\em extreme quantiles} $F^{-1}(1-p)$ for values $p$ very close to
zero. If the tail index {\em or} the scale parameter in a \ds\ of type
\eqref{eq:3} change inside a sample, then it is likely that the
extreme quantiles change as well. We will test for a change of tail
index or scale parameter in this indirect way.

@@

\subsection{An utility based argument for equal tail indices}\label{sec:3}
In this section we give an argument based on economic theory which  favors equality of  tail indices for return series.
We follow an approach by Routledge and Zin \cite{routledge2010generalized} who introduced the notion of
Generalized Disappointment Aversion (GDA). We consider the risky payoff $C$ of an investor and assume that
it has a continuous \ds\ on $(0,\infty)$ with \df\ $F_C$. Let $u$ be a utility \fct\ assumed to be increasing and concave on $(0,\infty)$.
Following Routledge and Zin \cite{routledge2010generalized},
the utility of an agent with GDA preferences is given by
\beao% \label{eq:xxie0}
  \wt u&=& \E [u(C)] - b\, \int_{0}^{\delta v}
  \big[ u(\delta \,v) - u(x) \big] F_C(dx)\nonumber\\&=& \E u(C) - b \E\big[\big(u(\delta \,v) - u(C)\big)\1( C < \delta \,v)\big]\,,
\eeao
where $\delta $ and $v$ are positive constants, and $b\ge 0$.
{\blue 
 Here $v$ can be thought of as the certainty payoff equivalent to the
 risky payoff $C$; $\delta$ tunes the threshold of {\em
   disappointment} in proportion to $v$; $b$ determines the extra
 weight given to the expected return of $C$ when $C$ is below the
 disappoinment threshold $\delta v$.
 If $b=0$, preferences are the classical expected utility. If
 $\delta=1$ and $b>0$ preferences follow Gul's \cite{gul:1991} 
disappointment aversion which were generalized by Routledge and Zin
\cite{routledge2010generalized}.
}
An agent guided by the utility function $u$ will seek to maximize the
\fct al $\wt u$.
Routledge and Zin assumed a power-law utility function 
\beam\label{eq:u}
u(x)=\frac{1}{1 - \gamma}\,x^{1 - \gamma}\,,\quad \gamma>0\,. 
\eeam
In practice it is more often to observe $\gamma > 1$ {\blue Let's
consult Casper for references.} Therefore we adopt a more convenient
notation in this case
\beao
u(x)=-\frac{1}{\xi}\,x^{-\xi}\,,\quad \xi>0\,. 
\eeao

@@

\subsection{Pareto-distributed returns}\label{sec:pareto_tail}
We consider the case when $X$ has a two-sided Pareto \ds\ given by
\beam\label{eq:pareto}
  F_X(x) = \left\{
  \begin{array}{ll}
    p \left(
    {K \over K - x}
    \right)^\alpha & x \leq 0 \,,\\
    1 - (1 - p) \left(
    {K' \over K' + x}
    \right)^\beta & x > 0\,,
  \end{array}
  \right.
\eeam
where 
$\alpha, \beta > 1$, $0 < p < 1$ and $K,K' > 0$. Then we have
\beam\label{eq:xxie1.0}
 \wt u(F_X,\phi)\nonumber
  &=&
  \alpha\, K^\alpha\,  p\,
  \int_{-\infty}^0
  u\big( (1 - \phi) \ex^r + \phi \ex^x \big)\,
  {1 + b \over (K - x)^{\alpha + 1}} \,dx\\
    &&+
  \beta \,(K')^\beta\, (1 - p)
  \int_{0}^\infty
  u\big( (1 - \phi) \ex^r + \phi\, \ex^x \big)\,
  {1 + b \1(x < q) \over (K' + x)^{\beta + 1}} \,dx 
  - b \,u(\delta v) \,F_X(q) \,.\label{eq:pp}
\eeam
{\blue
\begin{remark}\label{thrm:I}
 Assume the two-sided Pareto model \eqref{eq:pp} and that the utility
 \fct\ $u$ is increasing and differentiable. Then $\frac{\pd \wt
   u_{\rm max}}{\pd \alpha} > 0$ and $\frac{\pd \wt u_{\rm max}}{\pd
   K} < 0$.
\end{remark}
}
The proof is given in Appendix~\ref{sec:thrmI_proof}. 
\par
We illustrate the results of remark~\ref{thrm:I} in Figure~\ref{fig:preference_pareto} for the power-utility
\fct\ \eqref{eq:u}.
For this reason, we replace $\phi$ in the first integral in \eqref{eq:pp} by the optimal $\hat\phi$. The resulting
quantity is denoted by $I(\alpha, K)$.
\par
We conclude that
$\wt u_{\rm max}$ increases with $\alpha$ and decreases with $K$. Therefore there is a curve of
equal preference on the $(\alpha, K)$-plane. Moving along this
curve in the direction of increasing $\alpha$, one expects the values
of $K$ to increase too, i.e., the estimated values of $\alpha$ and $K$
should appear positively dependent. This is indeed the case for some
real return data, e.g. the ``Energy'', ``Consumer Staples'' and
``Information Technology'' sectors of the S\&P 500 index; see 
Figure \ref{fig:sectors_parameters}.
\par
The value of $\hat\phi$ can be calculated
by numerical integration and optimization with respect to $\phi$ for given values of $K, K', \alpha, \beta$. 
This is shown in 
Figure \ref{fig:phi_hat_pareto} for the power utility \fct\ \eqref{eq:u}.
\begin{figure}
  \begin{minipage}{0.5\linewidth}
    \includegraphics[width=\textwidth]{phi_hat_pareto5e-1.pdf}    
  \end{minipage}\hfill
  \begin{minipage}{0.5\linewidth}
    \includegraphics[width=\textwidth]{phi_hat_pareto4.pdf}
  \end{minipage}
  \caption{The graphs show $\hat\phi$, the optimal equity allocation
    as a \fct\ of $\alpha$ and $K$ in the two-sided Pareto model
    \eqref{eq:pareto} with $K=K'$, $\alpha=\beta$. We choose
    the utility $u$ from \eqref{eq:u}
    function $u(x) = -x^{-\xi}/\xi$ for $\xi = 1/2$ (left) and $\xi = 4$
    (right).
  }
  \label{fig:phi_hat_pareto}
\end{figure}

\begin{minipage}{0.5\linewidth}
  \includegraphics[width=\textwidth]{power_utilities.png}
\end{minipage}\hfill
\begin{minipage}{0.42\textwidth}
  In this graph we show two power-utility functions $u$. We see that
  $u(x)$ grows slower and saturates later for small $\xi$ than for
  large $\xi$. The latter case represents an agent who is more
  tolerant about low consumption and seeks wealth more
  aggressively. In other words, he is less risk-averse than one with a
  larger $\xi$. Such an agent will therefore invest more heavily in
  equity, as shown in Figure~\ref{fig:phi_hat_pareto}.
\end{minipage}

@@

\subsection{Student's t-distributed equity return}
It is a common practice to use Student's t-distribution to model the
stationary distribution of equity returns. So it is of interest to
find out what implications this distribution has when it is combined
with the PDA preference. Formally we assume
\[
f(x; \nu) = c(\nu) \left(
  1 + {x^2 \over \nu}
\right)^{-(\nu + 1)/2}
\]
where $\nu > 1$ and
\[
c(\nu) = {
  \Gamma({\nu + 1 \over 2})
  \over
  \Gamma(\nu/2) \sqrt{\nu \pi}
}
\]
We can write $\wt u(F, \phi)$ as
\begin{eqnarray}
  && \wt u(F, \phi) \nonumber \\
  &=&
  \int_{0}^{\infty}
  \left[
    u(C(x, \phi)) + (1 + b)u(C(-x, \phi))
  \right]
  f(x, \nu) dx \nonumber \\
  &&
  + b \int_{0}^{q}
  u(C(x, \phi))
  f(x, \nu) dx - b u(C(q, \phi)) F(q, \nu)
  \label{eq:t1}
\end{eqnarray}
where $C(\cdot)$ is defined in \eqref{eq:xxie1}.

@@

First notice that when $b = 0$, $\wt u(F, \phi)$ reduces to
$\E u(C(X, \phi))$. Only the first integral of \eqref{eq:t1} remains.
Its integrand becomes $\ln[C(x, \phi) C(-x, \phi)]$.
Now observe $\ln(C(x, \phi)C(-x, \phi))$ is an increasing function of
$x$:
\begin{eqnarray*}
  {\pd C(x, \phi) C(-x, \phi) \over \pd x}
  &=&
  \phi e^r (1 - \phi) (e^x - e^{-x}) > 0
\end{eqnarray*}
Thus when $b = 0$, by 1st case of remark \ref{remark:I} and lemma
\ref{lemma:II},
${\pd \wt u(F, \phi)) \over \pd \nu} < 0$ for all $\phi$. Hence all
investors guided by this utility function would be seeking the
smallest $\nu$, i.e. the heaviest tail in the market. Here note that
Student's t is a symmetric distribution; a heavier tail implies not
only greater down-side risk but also greater up-side potential.

In fact, even when $b > 0$, the same monotonicity property holds. This
is revealed by numerically maximizing $\wt u(F, \phi)$ with
respect to $\phi$ for a sequence of selected values of $\nu$. The
results are shown in figure~\ref{fig:phi_hat_U}.
\begin{figure}[htb!]
  % \begin{subfigure}[b]{0.5\linewidth}
  %   \includegraphics[width=\textwidth]{phi_hat_b_t_log.pdf}
  %   \caption{$\hat\phi(\nu)$ of log-utility function}
  %   \label{fig:phi_hat_t_log}
  % \end{subfigure}
  % \begin{subfigure}[b]{0.5\linewidth}
  %   \includegraphics[width=\textwidth]{U_b_t_log.pdf}
  %   \caption{$G(\nu)$ of log-utility function}
  %   \label{fig:U_b_t_log}
  % \end{subfigure}
  % \begin{subfigure}[b]{0.5\linewidth}
  %   \includegraphics[width=\textwidth]{phi_hat_b_t_power.pdf}
  %   \caption{$\hat\phi(\nu)$ of power-utility function}
  %   \label{fig:phi_hat_t_power}
  % \end{subfigure}
  % \begin{subfigure}[b]{0.5\linewidth}
  %   \includegraphics[width=\textwidth]{U_b_t_power.pdf}
  %   \caption{$G(\nu)$ of power-utility function}
  %   \label{fig:U_b_t_power}
  % \end{subfigure}
  \begin{minipage}{0.5\linewidth}
    \includegraphics[width=\textwidth]{phi_hat_b_t_log.pdf}
  \end{minipage}\hfill
  \begin{minipage}{0.5\linewidth}
    \includegraphics[width=\textwidth]{U_b_t_log.pdf}
  \end{minipage}
  \begin{minipage}{0.5\linewidth}
    \includegraphics[width=\textwidth]{phi_hat_b_t_power.pdf}
  \end{minipage}\hfill
  \begin{minipage}{0.5\linewidth}
    \includegraphics[width=\textwidth]{U_b_t_power.pdf}
  \end{minipage}
  \caption{$\hat\phi$ and $G(\nu)$
    \label{fig:phi_hat_U}
  }
\end{figure}

%% However, this is not necessarily the case when $b > 0$. For example,
%% consider the case when $q = 0$ and $\hat\phi = 1$. This may be
%% viewed an approximation to the situation where the interest rate $r$ is
%% vanishingly low while the investor is more conservative than
%% greedy. In this case we have
%% \begin{eqnarray*}
%%   \wt u(F, \phi) = \int_{0}^{\infty} e^{-b x} f(x, \nu) dx
%% \end{eqnarray*}
%% Thus by the 1st case of lemma \ref{lemma:I} and lemma \ref{lemma:II},
%% $\opd{\nu}\wt u(F, \phi) > 0$. In this extreme case, the investor
%% would rather seek the largest $\nu$ or in other words, the lightest
%% tail in the market.

\subsubsection{power-utility function}
In this section we consider the case when the utility function
$u(\cdot)$ takes the form
\[
u(x) = -{1 \over \xi} x^{-\xi}
\]
One can see from figure \ref{fig:phi_hat_U_power} that both $\hat
\phi$ and $G(\nu)$ are monotonically decreasing with $\nu$ as in the
case of a log-utility function.
\begin{figure}[htb!]
  \begin{minipage}{0.5\linewidth}
    \includegraphics[width=\textwidth]{phi_hat_b_t_power.pdf}
  \end{minipage}\hfill
  \begin{minipage}{0.5\linewidth}
    \includegraphics[width=\textwidth]{U_b_t_power.pdf}
  \end{minipage}
  \caption{$\hat\phi$ and $G(\nu)$ when the utility
    function is $u(C) = -{2 \over \sqrt{C}}$
  }
  \label{fig:phi_hat_U_power}
\end{figure}
